Exploring the Pareto front of multi-objective COVID-19 mitigation policies using reinforcement learning

### 2023-03-31 paper/main.tex
pieter comments

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,bm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{cleveref}       % smart cross-referencing
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{bm}             % bold fonts for math symbols
\usepackage[misc]{ifsym}    % collection of symbols
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}          % force figure positioning in text
\usepackage{chngcntr}       % counters per section
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{comment}
\usepgfplotslibrary{dateplot, groupplots, fillbetween}
\pgfplotsset{compat=newest}
\usepackage{plotpolicies}   % custom package to plot policy-executions

@@

The main disadvantage of this metric is that we need the true Pareto front to compute it, which is unknown for our MOMDP (see Section~\ref{sec:sars-cov2-momdp}). To still gain insights from our learned policies, we approximate the true Pareto front using the non-dominated policies across all runs.

@@

Figure~\ref{fig:paretofrontmetrics} shows a visual representation of the hypervolume and $\varepsilon$ metrics in two dimensions.

@@

A visual representation of the model is depicted in Fig~\ref{fig:ode_mdel}. It shows the different compartments, as well as the flow rates at which individuals move from one compartment to another.

@@

We denote $\hat{C}_t$ the social contact matrix at timestep $t$, resulting from the reduction function $\hat{C}$. Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in SI).

@@

    \caption{Schematic diagram of the compartmental model for SARS-CoV-2 presented by \citet{abrams2021modelling} which is used to derive the MOMDP.}
    \label{fig:ode_mdel}
\end{figure}
In order to apply multi-objective reinforcement learning, we construct the \momdpname\ MOMDP based on the epidemiological model introduced in Section~\ref{sec:sars-cov2-momdp_model} and graphically depicted in Figure~\ref{fig:ode_mdel}. Moreover, we consider a finite-horizon setting where we simulate the compartment model for a fixed number of weeks.

\paragraph{Action-space:} Our actions concern the installment of a social contact matrix with a particular reduction resulting from the reduction function $\hat{C}$ (see Section~\ref{sec:interventions}). To this end, we use the proportional reduction parameters $p_w, p_s, p_l$ defined in Section~\ref{sec:interventions}. Thus, each $\mdpaction \in \mdpactionspace$ is a 3-dimensional continuous vector in $[0,1]^3$ (i.e., $\mdpaction = [p_{w}, p_{s}, p_{l}]$) which impacts the social contact matrix according to Equation~\ref{eq:scm}.

\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M(\mdpstate_m' \mid \mdpstate_m, \hat{C}, p_{w}, p_{s}, p_{l})$ (see SI for details on $M$), where $\hat{C}(p_{w}, p_{s}, p_{l})$ returns the currently installed social contact matrix, that progresses the epidemiological model in one timestep.
We use this function as the transition function in \momdpname.

@@

We parameterize the epidemiological model using the mean of the posteriors as specified by \citet{abrams2021modelling} (details in SI).

@@

\begin{equation}
    \mdprewardfn_{\text{SB}}(\mdpstate,\mdpaction,\mdpstate') = \sum_{i=1}^{\agegroups}\sum_{j=1}^{\agegroups}(\hat{C}-C)_{ij}S_i(\mdpstate)S_j(\mdpstate) + \sum_{i=1}^{\agegroups}\sum_{j=1}^{\agegroups}(\hat{C}-C)_{ij}R_i(\mdpstate)R_j(\mdpstate),
\end{equation}
where $S_k(\mdpstate)$ represents the number of susceptible individuals in age group $k$ in state $\mdpstate$ and $R_k$ represents the number of recovered individuals in age group $k$ in state $\mdpstate$. In Section~\ref{sec:experiments}, we optimize PCN on two different variants for the multi-objective reward function: $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to study the impact of these distinct attack rate quantities.

@@

Concretely, the PCN algorithm implements an iterative process that repeatedly executes 3 main steps:
\begin{enumerate}
    \item Use the current policy to interact with the environment and generate trajectories (see Section~\ref{sec:pcn-policy-exploration}). Initially, use a random policy to fill the dataset with random trajectories.
    \item Update the dataset to incorporate these new trajectories and prune the least interesting trajectories in terms of returns (see Section~\ref{sec:pcn-building-dataset}).
    \item Update the policy by training it on the updated dataset (see Section~\ref{sec:pcn-continuous}).
\end{enumerate}

@@

While the original PCN algorithm is designed for MOMDPs with discrete actions-spaces, the problem we tackle (see Section~\ref{sec:sars-cov2-momdp}) is defined in terms of a continuous action-space. We thus extend PCN to the continuous action-space setting.
We change the output of the neural network such that there is a single output value for each dimension of the action-space. Since the actions should be bound in the domain of possible actions ($[0,1]$ in the case of \momdpname, see Section~\ref{sec:sars-cov2-momdp}), we apply a tanh non-linearity function on each output.
As such, we have a regression problem instead of a classification problem, as the labeled dataset now uses continuous labels $y = \mdpaction_t$ instead of categories. We thus use a Mean Squared Error (MSE) loss to update the policy:

@@

Although we use a stochastic model to cope with the uncertainty of the outcome of the outbreak, it is possible to deterministically evaluate the set of ordinary differential equations that define the model. We assess the validity of our approach by executing PCN on this deterministic variant, and observing similar performance as with the \momdpname\ model. \todo{Results are available in SI}.

@@

To conduct this analysis, we apply our extension of PCN for continuous action-spaces on the \momdpname\ model. As explained in Section~\ref{sec:pcn-stochastic}, we extend PCN for environments with stochastic transitions.

@@

All experiments are averaged over 5 runs. The hyper-parameters and the neural network architecture can be found in the SI.

@@

\begin{figure}
    \centering
    \input{fig/arh_binomial_both}
    \caption{The Pareto front of policies discovered by PCN using \momdpname, showing the different compromises between the number of hospitalizations and the number of lost contacts. On the right, we show, for each budget setting (colored, subscript indicates budget) the coverage set learned by the best performing run. On the left, we show an interpolated average of the coverage sets learned by the different runs, with the shaded regions corresponding to the standard deviation. For comparison, the basline is displayed on both plots (in black). As the budget increases, so does the size of the coverage set learnt by PCN. Changes are most noticeable in the less restrictive trade-offs in terms of social burden.}
    \label{fig:pf-binomal-arh-budgets}
\end{figure}

@@

We learn a coverage set (see Figure~\ref{fig:pf-binomal-arh-budgets}) that ranges from imposing minimal restrictions to enforcing many restrictions. In Figure~\ref{fig:pf-binomal-arh-budgets}, we display on the right the coverage set of the best-performing run in terms of hypervolume, for each budget setting. On the left, we show an interpolated average of the coverage sets learned by the different runs.

Regardless of the imposed budget, we notice that the coverage sets discovered by PCN almost completely dominate the coverage set of the baseline, demonstrating that there are better alternatives to the fixed policies. This is most evident in the compromising policies, where one has to carefully choose when to remove social restrictions while at the same time minimizing the impact on daily new hospitalizations. In these scenarios, PCN learns policies that drastically reduce the total number of new hospitalizations (e.g., more than $20000$) for the same social burden. We analyse the executions of such policies in Figure~\ref{fig:policy-execution} (middle plot), that shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the curve of the epidemic is flattened and gradually decreases over time.

Interestingly, we notice that the most restrictive policy (i.e. the one that prioritizes hospitalizations over social burden, see Figure~\ref{fig:policy-execution}, bottom plot) still starts to gradually increase $p_w$ and $p_l$ from the end of July onward. This is because by then, the epidemic has mostly faded out, and it is safe to reduce social restrictions. The timing of this reduction is important as reducing restrictions too soon can lead to a new wave. PCN learns the impact of its decisions over time, and correctly infers the timing at which restrictions can be safely lifted.

@@

Figure~\ref{fig:pf-binomal-arh-budgets} demonstrates that the budget impacts the learnt coverage set. In general, an increase of budget is associated with an increasingly better coverage set, as policies learned using a higher budget dominate the ones learned with a lower budget. This is to be expected, as a higher budget give the agent more freedom to change its actions as the epidemic progresses. However, we also notice that this increase of coverage sets is not necessarily gradual. Between the run with a budget of 2 and budget of 3, the difference in coverage set seems marginal. \todo{insight}.

Moreover, we observe that the difference is concentrated around the less restrictive policies in terms of social burden. We hypothesise that this region contains the most complex policies, as these try to maintain as much social freedom as possible, while containing the number of hospitalisations. In these cases, the timing of the actions greatly correlate with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control about managing this timing. To confirm this, we select, for each budget setting, the solution where the difference in performance is most noticeable, corresponding to the solutions with a total number of hospitalisations around $80000$. We plot the execution of the corresponding policies in Figure~\ref{fig:pe-binomial-arh-budgets} and analyse their impact in terms of social burden. First, we observe that the lower-budget policies are unable to reduce the social restrictions past the peak of the epidemic. In contrast, the setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave. Second, we note that the policy with a budget of 5 ressembles the execution of the one without restrictions. However, due to its budget, the policy is unable to progressively reduce the restrictions and instead resorts to a halfway compromise. \todo{corresponding SB numbers.} Compared to this specific region of the coverage set, the difference in performance between different budget settings seems marginal around the extremas. At the extremas, the policies are less complex (e.g., business-as-usual, resulting in the same action executed throughout the episode) and are thus less impacted by the budget restrictions.

Finally, we observe that, while the extremas deliver similar trade-offs for any of the chosen budgets, these trade-offs differ for the setting without budget restrictions. Indeed, in this setting, PCN does not learn the most extreme policies with respect to restrictions, even though there are no constraints on the action-set. As explained in Section~\ref{sec:pcn-policy-exploration}, PCN searches for increasingly better solutions using a stochastic policy. Thus, at every timestep, the action can change compared to the previous one. Continuously outputting the same action (e.g., no social restrictions) becomes a complicated task. In comparison, for the settings with a limited budget, the action stays the same as the previous one once the budget has been spent. As such, it is easier to learn the most extreme policies. Thus, in the specific case where we impose no budget, the freedom of action actually hinders PCN's search, for certain regions of the rweard-space.

@@

Even with the $\mdprewardfn_\text{ARI}$, the increased budget shows an increase in hypervolume in terms of hospitalisations. Moreover, those hypervolumes are close to the ones trained on $\mdprewardfn_\text{ARH}$, indicating that their coverage sets are similar. However, regardless of the imposed budget, the hypervolumes are slightly worse. This is to be expected, since those experiments are not directly optimising on $\mdprewardfn_\text{ARH}$. We make similar conclusions for $I_\varepsilon$. For budgets 2, 3 and 5, the difference between the worst-performing policy for the $\mdprewardfn_\text{ARI}$ variant and the $\mdprewardfn_\text{ARH}$ is less than $0.01$, indicating less than $1\%$ difference in return values between the two variants when comparing their worst-performing policy. As an exception, we notice that PCN without budget restrictions results in better performance across every metric for the $\mdprewardfn_\text{ARI}$ variant. Still, due to the high standard deviation of the no-budget, $\mdprewardfn_\text{ARH}$ setting, we do not believe this difference is meaningfull. \todo{some kind of meaningful conclusion}

@@

The dataset of trajectories that PCN is trained on is pruned over time to keep only the most relevant trajectories. The returns of these trajectories are used in Figure~\ref{fig:pf-binomal-arh-budgets} to show the learned coverage set. Each of these returns can be used as desired return for policy execution. We now assess the robustness of the executed policies, by comparing the return obtained after executing the policy with the corresponding target return. For each run, we execute the policy $10$ times and compute the $I_\varepsilon$ and $I_{\varepsilon-mean}$ metrics with respect to the coverage set learned during the run. We show that the executed policies reliably obtain returns that are similar to the desired return used to condition PCN.

@@

To conclude, we show that multi-objective reinforcement learning can be used to learn a wide set of high-quality policies on real-world problems, providing the decision maker with insightful and diverse alternatives, and showing the impact of taking extreme measures.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}
\appendix
\counterwithin{figure}{section}
\counterwithin{table}{section}
\counterwithin{equation}{section}

\section{Compartmental model for SARS-CoV-2}
In this work we utilise the compartmental model proposed by \citet{abrams2021modelling} and extend the model to a multi-objective Markov decision process (MOMDP). The MOMDP used in this work is outlined in the main paper. However, we describe the details of the underlying compartmental model used to model the spread of COVID-19 below.

\begin{figure}[h]
    \centering
     \begin{tikzpicture}
     \draw (0.0pt, 0.0pt)node[fill=green!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](0){$\textbf{S}$};
     \draw (60.0pt, 0.0pt)node[fill=orange!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](1){$\textbf{E}$};
     \draw (120.0pt, 0.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](2){$\textbf{I}^{presym}$};
     \draw (180.0pt, 10.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](3){$\textbf{I}^{asym}$};
     \draw (170.0pt, -30.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](4){$\textbf{I}^{mild}$};
     \draw (170.0pt, -70.0pt)node[fill=red!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](5){$\textbf{I}^{sev}$};
     \draw (230.0pt, -50.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](6){$\textbf{I}^{hosp}$};
     \draw (230.0pt, -90.0pt)node[fill=blue!20, thick, minimum height=0.6cm,minimum width=0.6cm, draw](7){$\textbf{I}^{icu}$};
     \draw (300.0pt, -50.0pt)node[fill=orange!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](8){$\textbf{R}$};
     \draw (300.0pt, -90.0pt)node[fill=purple!50, thick, minimum height=0.6cm,minimum width=0.6cm, draw](9){$\textbf{D}$};

     %\path[->] (6) edge[dotted, bend right = 50]
     \path[->] (0) edge[] node[above]{$\boldsymbol{\lambda}(t)$} (1);
     \path[->] (1) edge[] node[above]{$\gamma$} (2);
     \path[->] (2) edge[] node[below]{$\textbf{p}\theta$} (3);
     \path[->] (2) edge[] node[near end, below left]{$(1-\textbf{p})\theta$} (4);
     \path[->] (4) edge[] node[left]{$\boldsymbol{\psi}$} (5);
     \path[->] (5) edge[] node[near end, below left]{$(1-\phi_{1})\boldsymbol{\omega}$} (7);
     \path[->] (5) edge[] node[above]{$\phi_{1}\boldsymbol{\omega}$} (6);
     \path[->] (7) edge[] node[near end, below right]{$\boldsymbol{\delta}_{4}$} (8);
     \path[->] (7) edge[] node[above]{$\boldsymbol{\tau}_{2}$} (9);
     \path[->] (6) edge[] node[above]{$\boldsymbol{\delta}_{3}$} (8);
     \path[->] (6) edge[] node[near start, below left]{$\boldsymbol{\tau}_{1}$} (9);
     \path[->] (3) edge[bend left = 30] node[above right]{$\delta_{1}$} (8);
     \path[->] (4) edge[bend left = 30] node[above left]{$\boldsymbol{\delta}_{2}$} (8);

    \end{tikzpicture}

    \caption{Schematic diagram of the compartmental model for SARS-CoV-2 presented by \citet{abrams2021modelling} which is used to derive the MOMDP.
    }
    \label{fig:ode_mdel}
\end{figure}

We utilise an adapted SEIR mathematical compartmental model proposed by Abrams et al. to contruct the MOBelCov MOMDP with deterministic and stochastic state transitions. In this model members of the population are susceptible to infection when they are in compartment $\textbf{S}$\footnote{Boldface vector notation is used to denote the multiple age groups for each compartment.}. If an individual comes into contact with an infections individual then they move to the exposed compartment, $\textbf{E}$, at a time specific rate, $\boldsymbol{\lambda} (t)$. After a period of time an exposed individual becomes infectious, where they move to the pre-symptomatic compartment, $\textbf{I}^{presym}$, at rate $\gamma$. Once infected, individuals develop mild symptoms, $\textbf{I}^{mild}$ , with probability $1 - \textbf{p}$ or do not develop any symptoms, $\textbf{I}^{asym}$, with probability $\textbf{p}$, where asymptomatic individuals recover, $\textbf{R}$, at rate $\delta_{1}$. Individuals who experience symptoms can suffer from a mild infection, $\textbf{I}^{mild}$, and recover at rate $\boldsymbol{\delta}_{2}$, or they suffer from a more severe infection, $\textbf{I}^{sev}$, at a rate $\boldsymbol{\psi}$. Individuals with a severe infection are then transferred to a hospital for treatment, $\textbf{I}^{hosp}$ with probability $\phi_{1}$. However, some individuals become critically ill and are transferred directly to the intensive care unit (ICU) with probability $1 - \psi_{1}$. Individuals in the hospital, $\textbf{I}^{hosp}$ and $\textbf{I}^{icu}$, recover at rate $\boldsymbol{\delta}_{3}$ or $\boldsymbol{\delta}_{4}$ and die at rate $\boldsymbol{\tau}_{3}$ and $\boldsymbol{\tau}_{4}$ respectively. Figure \ref{fig:ode_mdel} outlines the compartmental model defined by Abrams et al.~\cite{abrams2021modelling}.

\subsection{Deterministic Compartmental Model}
\label{sec:deterministic_comp_model}
The flows of the deterministic model are defined by a set of ordinary differential equations, which are outlined as follows:

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsmath,bm}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{cleveref}       % smart cross-referencing
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{bm}             % bold fonts for math symbols
\usepackage[misc]{ifsym}    % collection of symbols
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}          % force figure positioning in text
\usepackage{chngcntr}       % counters per section
\usepackage{tikz}
\usepackage{pgfplots,etoolbox}
\usepackage{comment}
\usetikzlibrary{external}
\usepgfplotslibrary{dateplot, groupplots, fillbetween}
\pgfplotsset{compat=newest}
\tikzexternalize[prefix=cache/]

\usepackage{plotpolicies}   % custom package to plot policy-executions

@@

The main disadvantage of this metric is that we need the true Pareto front to compute it, which is unknown for our MOMDP (see Sec.~\ref{sec:sars-cov2-momdp}). To still gain insights from our learned policies, we approximate the true Pareto front using the non-dominated policies across all runs.

@@

Fig.~\ref{fig:paretofrontmetrics} shows a visual representation of the hypervolume and $\varepsilon$ metrics in two dimensions.

@@

A visual representation of the model is depicted in Fig~\ref{fig:compartment_model}. It shows the different compartments, as well as the flow rates at which individuals move from one compartment to another.

@@

We denote $\hat{C}_t$ the social contact matrix at timestep $t$, resulting from the reduction function $\hat{C}$. Secondly, we assume that compliance to the interventions is gradual and model this using a logistic compliance function (see details in Sec.~\ref{sec:compliance-function}).

@@

    \caption{Schematic diagram of the compartmental model for SARS-CoV-2 presented by \citet{abrams2021modelling} which is used to derive the MOMDP.}
    \label{fig:compartment_model}
\end{figure}
In order to apply multi-objective reinforcement learning, we construct the \momdpname\ MOMDP based on the epidemiological model introduced in Sec.~\ref{sec:sars-cov2-momdp_model} and graphically depicted in Fig.~\ref{fig:compartment_model}. Moreover, we consider a finite-horizon setting where we simulate the compartment model for a fixed number of weeks.

\paragraph{Action-space:} Our actions concern the installment of a social contact matrix with a particular reduction resulting from the reduction function $\hat{C}$ (see Sec.~\ref{sec:interventions}). To this end, we use the proportional reduction parameters $p_w, p_s, p_l$ defined in Sec.~\ref{sec:interventions}. Thus, each $\mdpaction \in \mdpactionspace$ is a 3-dimensional continuous vector in $[0,1]^3$ (i.e., $\mdpaction = [p_{w}, p_{s}, p_{l}]$) which impacts the social contact matrix according to Eq.~\ref{eq:scm}.

\paragraph{Transition function:} The model defined by \citet{abrams2021modelling} utilises a model transition probability $M(\mdpstate_m' \mid \mdpstate_m, \hat{C}, p_{w}, p_{s}, p_{l})$ (see Sec.~\ref{sec:binomial-model} for details on $M$), where $\hat{C}(p_{w}, p_{s}, p_{l})$ returns the currently installed social contact matrix, that progresses the epidemiological model in one timestep.
We use this function as the transition function in \momdpname.

@@

We parameterize the epidemiological model using the mean of the posteriors as specified by \citet{abrams2021modelling} (details in Sec.~\ref{sec:compartment-model-parameters}).

@@

\begin{equation}
    \mdprewardfn_{\text{SB}}(\mdpstate,\mdpaction,\mdpstate') = \sum_{i=1}^{\agegroups}\sum_{j=1}^{\agegroups}(\hat{C}-C)_{ij}S_i(\mdpstate)S_j(\mdpstate) + \sum_{i=1}^{\agegroups}\sum_{j=1}^{\agegroups}(\hat{C}-C)_{ij}R_i(\mdpstate)R_j(\mdpstate),
\end{equation}
where $S_k(\mdpstate)$ represents the number of susceptible individuals in age group $k$ in state $\mdpstate$ and $R_k$ represents the number of recovered individuals in age group $k$ in state $\mdpstate$. In Sec.~\ref{sec:experiments}, we optimize PCN on two different variants for the multi-objective reward function: $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to study the impact of these distinct attack rate quantities.

@@

Concretely, the PCN algorithm implements an iterative process that repeatedly executes 3 main steps:
\begin{enumerate}
    \item Use the current policy to interact with the environment and generate trajectories (see Sec.~\ref{sec:pcn-policy-exploration}). Initially, use a random policy to fill the dataset with random trajectories.
    \item Update the dataset to incorporate these new trajectories and prune the least interesting trajectories in terms of returns (see Sec.~\ref{sec:pcn-building-dataset}).
    \item Update the policy by training it on the updated dataset (see Sec.~\ref{sec:pcn-continuous}).
\end{enumerate}

@@

While the original PCN algorithm is designed for MOMDPs with discrete actions-spaces, the problem we tackle (see Sec.~\ref{sec:sars-cov2-momdp}) is defined in terms of a continuous action-space. We thus extend PCN to the continuous action-space setting.
We change the output of the neural network such that there is a single output value for each dimension of the action-space. Since the actions should be bound in the domain of possible actions ($[0,1]$ in the case of \momdpname, see Sec.~\ref{sec:sars-cov2-momdp}), we apply a tanh non-linearity function on each output.
As such, we have a regression problem instead of a classification problem, as the labeled dataset now uses continuous labels $y = \mdpaction_t$ instead of categories. We thus use a Mean Squared Error (MSE) loss to update the policy:

@@

Although we use a stochastic model to cope with the uncertainty of the outcome of the outbreak, it is possible to deterministically evaluate the set of ordinary differential equations that define the model. We assess the validity of our approach by executing PCN on this deterministic variant, and observing similar performance as with the \momdpname\ model. We show the results in Sec.~\ref{sec:cs-ode-vs-binomial}.

@@

To conduct this analysis, we apply our extension of PCN for continuous action-spaces on the \momdpname\ model. As explained in Sec.~\ref{sec:pcn-stochastic}, we extend PCN for environments with stochastic transitions.

@@

All experiments are averaged over 10 runs. The hyper-parameters and the neural network architecture can be found in Sec.s~\ref{sec:pcn-hyperparameters} and Sec.~\ref{sec:pcn-architecture}, respectively.

@@

\begin{figure}
    \centering
    \input{fig/arh_binomial_both}
    \caption{The Pareto front of policies discovered by PCN using \momdpname, showing the different compromises between the number of hospitalizations and the number of lost contacts. On the right, we show, for each budget setting (colored, subscript indicates budget) the coverage set learned by the best performing run. On the left, we show an interpolated average of the coverage sets learned by the different runs, with the shaded regions corresponding to the standard deviation. For comparison, the basline is displayed on both plots (in black). As the budget increases, so does the size of the coverage set learnt by PCN. Changes are most noticeable in the less restrictive trade-offs in terms of social burden.}
    \label{fig:pf-binomial-arh-budgets}
\end{figure}

@@

We learn a coverage set (see Fig.~\ref{fig:pf-binomial-arh-budgets}) that ranges from imposing minimal restrictions to enforcing many restrictions. In Fig.~\ref{fig:pf-binomial-arh-budgets}, we display on the right the coverage set of the best-performing run in terms of hypervolume, for each budget setting. On the left, we show an interpolated average of the coverage sets learned by the different runs.

Regardless of the imposed budget, we notice that the coverage sets discovered by PCN almost completely dominate the coverage set of the baseline, demonstrating that there are better alternatives to the fixed policies. This is most evident in the compromising policies, where one has to carefully choose when to remove social restrictions while at the same time minimizing the impact on daily new hospitalizations. In these scenarios, PCN learns policies that drastically reduce the total number of new hospitalizations (e.g., more than $20000$) for the same social burden. We analyse the executions of such policies in Fig.~\ref{fig:policy-execution} (middle plot), that shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the curve of the epidemic is flattened and gradually decreases over time.

Interestingly, we notice that the most restrictive policy (i.e. the one that prioritizes hospitalizations over social burden, see Fig.~\ref{fig:policy-execution}, bottom plot) still starts to gradually increase $p_w$ and $p_l$ from the end of July onward. This is because by then, the epidemic has mostly faded out, and it is safe to reduce social restrictions. The timing of this reduction is important as reducing restrictions too soon can lead to a new wave. PCN learns the impact of its decisions over time, and correctly infers the timing at which restrictions can be safely lifted.

@@

Fig.~\ref{fig:pf-binomial-arh-budgets} demonstrates that the budget impacts the learnt coverage set. In general, an increase of budget is associated with an increasingly better coverage set, as policies learned using a higher budget dominate the ones learned with a lower budget. This is to be expected, as a higher budget give the agent more freedom to change its actions as the epidemic progresses. However, we also notice that this increase of coverage sets is not necessarily gradual. % Between the run with a budget of 2 and budget of 3, the difference in coverage set seems marginal. \todo{insight}.

Moreover, we observe that the difference is concentrated around the less restrictive policies in terms of social burden. We hypothesise that this region contains the most complex policies, as these try to maintain as much social freedom as possible, while containing the number of hospitalisations. In these cases, the timing of the actions greatly correlate with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control about managing this timing. To confirm this, we select, for each budget setting, the solution where the difference in performance is most noticeable, corresponding to the solutions with a total number of hospitalisations around $80000$ (which is in the middle of the range of possible hospitalisations, as can be seen in Fig.~\ref{fig:pf-binomial-arh-budgets}). We plot the execution of the corresponding policies in Fig.~\ref{fig:pe-binomial-arh-budgets} and analyse their impact in terms of social burden. First, we observe that the lower-budget policies are unable to reduce the social restrictions past the peak of the epidemic. In contrast, the setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave. Second, we note that the policy with a budget of 5 ressembles the execution of the one without restrictions. However, due to its budget, the policy is unable to progressively reduce the restrictions and instead resorts to a halfway compromise. Compared to this specific region of the coverage set, the difference in performance between different budget settings seems marginal around the extremas. At the extremas, the policies are less complex (e.g., business-as-usual, resulting in the same action executed throughout the episode) and are thus less impacted by the budget restrictions.

Finally, we observe that, while the extremas deliver similar trade-offs for any of the chosen budgets, these trade-offs differ for the setting without budget restrictions. Indeed, in this setting, PCN does not learn the most extreme policies with respect to restrictions, even though there are no constraints on the action-set. As explained in Sec.~\ref{sec:pcn-policy-exploration}, PCN searches for increasingly better solutions using a stochastic policy. Thus, at every timestep, the action can change compared to the previous one. Continuously outputting the same action (e.g., no social restrictions) becomes a complicated task. In comparison, for the settings with a limited budget, the action stays the same as the previous one once the budget has been spent. As such, it is easier to learn the most extreme policies. Thus, in the specific case where we impose no budget, the freedom of action actually hinders PCN's search, for certain regions of the rweard-space.

@@

Even with the $\mdprewardfn_\text{ARI}$, the increased budget shows an increase in hypervolume in terms of hospitalisations. Moreover, those hypervolumes are close to the ones trained on $\mdprewardfn_\text{ARH}$, indicating that their coverage sets are similar. However, regardless of the imposed budget, the hypervolumes are slightly worse. This is to be expected, since those experiments are not directly optimising on $\mdprewardfn_\text{ARH}$. We make similar conclusions for $I_\varepsilon$. For budgets 2, 3 and 5, the difference between the worst-performing policy for the $\mdprewardfn_\text{ARI}$ variant and the $\mdprewardfn_\text{ARH}$ is less than $0.01$, indicating less than $1\%$ difference in return values between the two variants when comparing their worst-performing policy. As an exception, we notice that PCN without budget restrictions results in better performance across every metric for the $\mdprewardfn_\text{ARI}$ variant. Still, due to the high standard deviation of the no-budget, $\mdprewardfn_\text{ARH}$ setting, we do not believe this difference is meaningful. \todo{some kind of meaningful conclusion}

@@

The dataset of trajectories that PCN is trained on is pruned over time to keep only the most relevant trajectories. The returns of these trajectories are used in Fig.~\ref{fig:pf-binomial-arh-budgets} to show the learned coverage set. Each of these returns can be used as desired return for policy execution. We now assess the robustness of the executed policies, by comparing the return obtained after executing the policy with the corresponding target return. For each run, we execute the policy $10$ times and compute the $I_\varepsilon$ and $I_{\varepsilon-mean}$ metrics with respect to the coverage set learned during the run. We show that the executed policies reliably obtain returns that are similar to the desired return used to condition PCN.

@@

\todo{Although we use an age-structured compartment model, with social contact matrices to model social interactions, this remains a model that evaluates the progression of the pandemic as an aggregated process over the population. Individual-based models could provide more detailed and localized policies, potentially further improving the quality of possible trade-offs, and would provide an interesting avenue for future work. However, due to the computational cost of simulating such models, and the number of interactions required by reinforcement learning in general, this remains a challenging problem.}

\todo{Moreover, although we are able to interpret and analyse the obtained policies and their corresponding trade-offs, as we can plot the Pareto front for 2 objectives, this approach cannot be used for problems with more objectives. Defining interpretation and visualization tools for many-objective problems is necessary if we want to include additional objectives for our problem.}

\todo{Additionally, PCN is able to cope with model stochasticity, because the stochasticity is limited. In settings where the stochasticity is more pronounced, e.g., designing policies to control the initial outbreak of an epidemic, further methodological extensions are warranted.}

\todo{Finally, while we have focused on social burden as a trade-off for hospitalisations, our approach can be used for related problems, involving different compromises, such as balancing the number of lost schooldays with respect to new infections in schools \cite{torneri2021controlling}, contact tracing effort versus its effect \cite{willem2021impact}, the impact of antivirals on the epidemic versus resistance to antivirals \cite{torneri2020prospect}, or the cost of universal testing versus its impact on the epidemic \cite{libin2021assessing}.}

To conclude, we show that multi-objective reinforcement learning can be used to learn a wide set of high-quality policies on real-world problems, providing the decision maker with insightful and diverse alternatives, and showing the impact of taking extreme measures. \todo{This can be done for realistic settings, where we budget the number of changes of social reductions, such that the policies are practical and can be deployed in real-life scenarios.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\counterwithin{figure}{section}
\counterwithin{table}{section}
\counterwithin{equation}{section}

\section{Stochastic Compartmental Model}
\label{sec:binomial-model}
In this work we utilise the compartmental model proposed by \citet{abrams2021modelling} and extend the model to a multi-objective Markov decision process (MOMDP). Fig.~\ref{fig:compartment_model} shows a visual depiction of the compartment model. The flows of the deterministic model are defined by a set of ordienary differential equations, outlined in Sec.~\ref{sec:sars-cov2-momdp_model}.

Intervening in the spread of the virus by, for example, reducing social contacts or government interventions introduces uncertainty in the further course of the outbreak. Therefore, to understand how this uncertainty affect the spread of the disease we introduce a stochastic component model which can model the uncertainty generated by interventions in social contacts.

By formulating the set of differential equations defined in Sec.~\ref{sec:sars-cov2-momdp_model}, as a chain-binomial, we can obtain stochastic trajectories from this model \cite{bailey1975mathematical}. A chain-binomial model assumes a stochastic model where infected individuals are generated by some underlying probability distribution. For the stochastic model we consider a time interval $(t, t +h]$, where $h$ is defined as the length between two consecutive time points. Similar to \citet{abrams2021modelling}, in this work we set $h = \frac{1}{24}$. \citet{abrams2021modelling} define the set of differential equations as a chain binomial as follows:
